This is an example of article title

### 2025-11-01 paper/paper.tex
Fix a typo.
Fix metadata and giving the paper a careful read.

\newcommand{\term}[1]{\emph{#1}}
\newcommand{\e}{\ensuremath{\mathrm{e}}}
\newcommand{\im}{\ensuremath{\mathrm{i}}}
\newcommand{\LieGroupsVersion}{0.1.8}
%
% Define muted colors.
\definecolor{TolMutedBlue}{HTML}{332288}
\definecolor{TolMutedCyan}{HTML}{88CCEE}
\definecolor{TolMutedTeal}{HTML}{44AA99}
\definecolor{TolMutedGreen}{HTML}{117733}
\definecolor{TolMutedOlive}{HTML}{999933}
\definecolor{TolMutedSand}{HTML}{DDCC77}
\definecolor{TolMutedRose}{HTML}{CC6677}
\definecolor{TolMutedWine}{HTML}{882255}
\definecolor{TolMutedPurple}{HTML}{AA4499}
%
\hypersetup{colorlinks=true}
%
\begin{document}

@@

These spaces are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
They have applications in physics, robotics, stochastic processes, information geometry, and many other areas see~\cite{Chirikjian:2009,Chirikjian:2012},
but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
and their numerical aspects, for example when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998}.

@@

This paper provides an overview of the main features of \verb|LieGroups.jl| as of version \LieGroupsVersion{}.
After introducing the necessary mathematical background and notation in Section~\ref{sec:Notation},
we present the interface in Section~\ref{sec:Interface}, split into the Lie group, the Lie algebra, and group actions.
In Section~\ref{sec:LieGroups}, we present an overview of currently implemented Lie groups.
Finally, in Section~\ref{sec:Example}, we demonstrate how to get started and use \verb|LieGroups.jl|.

@@

which locally looks like \(\mathbb{R}^2\), think of charts in an atlas, but globally it is not.
Finally we denote the tangent space at a point \(p \in \mathcal{M}\) by \(T_p\mathcal{M}\). This can be thought of as all “velocities” (direction and speed) in which a curve can “pass through” a point. Formally it is set the equivalence classes of derivatives of smooth curves. Each tangent space $T_p\mathcal M$ is a $n$-dimensional vector space and we call the disjoint union of all tangent spaces
\begin{equation*}
    T\mathcal M = \dot\bigcup_{p \in \mathcal M} T_p\mathcal M
\end{equation*}
the \term{tangent bundle} of $\mathcal M$.
\\
As a group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\). Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{Note that the identification is not continuous.}, or, continuously, with the circle. % chktex 9ok
The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose (or an angle $-\alpha$). % -(-\pi) would be out of the range

\subsection{Tangent spaces and the Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
Then, using the differential (or push forward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$\footnote{\color{red} MB: TODO: does the reader know that a differential is? (RB) Sure I can expand on that, I wrote a sentence more further below to get from differential to Jacobian, so I can also add a sentence there as well.}, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields. \texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
\\
As a first example, consider $\mathcal G = (\mathbb R^n, +)$, where the tangent space at any point $g$ is again $\mathbb R^n$, especially at the identity $e=0$. We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
Here, a tangent vector $X$ induces the constraint vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$. Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \frak{g}$.

@@

Concerning a metric on the tangent spaces, manifolds are turned into Riemannian manifolds when they are equipped with a Riemannian metric \(\langle \cdot, \cdot \rangle_p\colon T_p\mathcal{M} \times T_p\mathcal{M} \to \mathbb{R}\) for each point \(p \in \mathcal{M}\) that smoothly varies with \(p\).
For Lie groups, we can use a single inner product \(\langle \cdot, \cdot \rangle\) on the Lie algebra \(\mathfrak{g}\) and use the change in representation as mentioned above to define
\begin{equation*}
\langle X, Y \rangle_g = \langle D\lambda_{g^{-1}}(g)[X], D\lambda_{g^{-1}}(g)[Y] \rangle_e
\end{equation*}
for all \(X, Y \in T_g\mathcal{G}\) and \(g \in \mathcal{G}\). Representing tangent vectors at arbitrary points \(g \in \mathcal{G}\) using the Lie algebra \(\mathfrak{g}\) yields here, that we can use the single inner product directly to evaluate this Riemannian metric. Hence representing tangent vectors using the Lie algebra is the default in the following.
\\
For the two examples above we obtain the Euclidean inner product \(\langle X, Y \rangle = X^{\mathrm{T}} Y\) for \(X, Y \in \mathbb{R}^n\) for \((\mathbb{R}^n, +)\) and the Frobenius inner product \(\langle X, Y \rangle = \mathrm{trace}(X^{\mathrm{T}} Y)\), \(X, Y \in \mathfrak{so}(n)\) can be used for \(\mathrm{SO}(n)\).

@@

\subsection{Group Actions} A \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\sigma\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\sigma(e, p) = p\) and \(\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)\).\footnote{This is the convention for left actions. Alternatively, right actions fulfil \(\sigma(g, \sigma(h, p)) = \sigma(h \cdot g, p)\).}
Informally a group action describes how elements of the Lie group \(\mathcal{G}\) “act on” points on the manifold \(\mathcal{M}\). As an example, think of the special orthogonal group \(\mathrm{SO}(3)\) acting on points on Euclidean space $\mathbb{R}^3$ “moving” them somewhere by rotating them around the origin. We obtain the group action \(\sigma\colon \mathrm{SO}(3) \times \mathbb{R}^3 \to \mathbb{R}^3\), \(\sigma(R, x) \coloneqq Rx\).

The same action can also be applied to points from the sphere \(\mathbb{S}^2\), resulting in a similar group action \(\sigma\colon \mathbb{SO}(3) \times \mathbb{S}^2 \to \mathbb{S}^2\), \(\sigma(R, p) \coloneqq Rp\).

@@

The interface follows the philosophy of \verb|ManifoldsBase.jl|, that the Lie group or algebra is the first argument and even a mutated argument comes second.
Similarly, if a function computes something like a new point on the Lie group or a tangent vector, there also exists a variant, that computes this in-place.
For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default representation, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types.

@@

In general, when a function within the interfaces of \verb|JuliaManifolds| returns a point, a tangent vector or any other mutable object, there always exist two variants of the function, one that allocates a new object and returns it, and one that takes an additional argument to write the result into. For example for a function \verb|f(G, g)| that returns a point on the Lie group \verb|G|, there also exists a variant \verb|f!(G, h, g)| that computes the result in-place of \verb|h|, possibly using the memory also for interims results. All methods are written such that the result is correct even if arguments \verb|h| and \verb|g| use the same memory. By default the allocating variant is implemented by allocating memory accordingly and calling the in-place variant. This default can be overriden, for example for performance reasons.

\subsubsection*{Topological functions}
Points and tangent vectors can be represented by the same Julia objects, although they are distinct from a topological viewpoint.%
\footnote{\texttt{LieGroups.jl} introduces an abstract type \texttt{AbstractLieGroupPoint <: AbstractManifoldPoint}, that is used to represent points on the Lie group.
This abstract supertype is not necessary, points can be represented by arrays or other types as well.}
Separate types are necessary when one Lie group has \emph{different representations} of points, that have to be distinguished. When there are different representations, it is recommended to introduce a point type for each representation and make the default one fall back to (just) using arrays.
Several functions in the interface return points or tangent vectors, where the type might not be known from the input argument. For such a function, for example again a function \verb|f(G)| a positional argument \verb|T| can be used to specify the desired return type, i.e.\ \verb|f(G, T)| returns a point or tangent vector of type \verb|T|.
The in-place variant does not require this, since in its signatgure of the form \verb|f!(G, g::T)| the type \verb|T| is known from the argument \verb|g|.

@@

In \verb|LieGroups.jl|, the group exponential function is given by \verb|exp(G,X)|%
%\footnote{in-place variant \texttt{exp!(G, g, X)}.}
.
If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\frak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|%
%\footnote{in-place variant \texttt{exp!(G, h, g, X)}.}
.
As a word of carefulness, note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map. The Riemannian exponential map \verb|exp(M, p, X)|
%\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.}
can be distinguished in that the first argument is a manifold, and the tangent vector \verb|X| has to be from the tangent space at the point \verb|p| on the manifold.
Here, again, to access the Riemannian exponential, one can use the base manifold of the Lie group, i.e.\ \verb|exp(base_manifold(G), g, X)| and have to make sure that \verb|X| is from the tangent space at \verb|g|.

@@

Similar to points on the Lie group, when representing elements $X \in \frak g$ from a Lie algebra, we do not type the general functions of the interface.
This allows to use either plain arrays or own structures to represent these in code.
\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}
Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.

@@

A central function on the Lie algebra is the Lie bracket $[\cdot, \cdot]\colon \frak g \times \frak g \to \frak g$, which is available as \verb|lie_bracket(|$\frak g$\verb|, X, Y)|.
% \footnote{in-place variant \texttt{lie\_bracket!($\frak g$, Z, X, Y)}.}
For more details on the Lie bracket see~\cite[Ch.~5]{HilgertNeeb:2012}.

\subsubsection*{Topological basics.}
Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\frak g$ \verb|= LieAlgebra(G)|.
To access the Lie group again, use \verb|base_lie_group(|$\frak g$\verb|)|. Similarly \verb|base_manifold(|$\frak g$\verb|)| returns the underlying manifold of the Lie group.\\
As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
A zero vector is generated via \verb|zero_vector(|$\frak g$\verb|)| for the default representation and \verb|zero_vector(|$\frak g$\verb|, T)| for a specific representation type \verb|T|%
%\footnote{the in-place variant \texttt{zero\_vector!($\frak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
.

The main topological function is to test the validity of a tangent vector \verb|is_vector(|$\frak g$\verb|, X; error=:none)| using the same \verb|error=| keyword as \verb|is_point| on the Lie group.

\subsubsection*{Vector space related functions}
For the following functions related to vector space features, using the Lie algebra $\frak g$ as first argument is equivalent to specifying the Lie Group $G$ and an arbitrary point $g \in \mathcal G$. This yields that the Lie group complies with the general interface for manifolds.

The inner product and norm on the Lie algebra are available as \verb|inner(|$\frak g$\verb|, X, Y)| and \verb|norm(|$\frak g$\verb|, X)|, respectively.
Furthermore, there are two functions to convert between a coordinate-free representation of $X$ as a tangent vector and its representation in coordinates of a basis.

Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\frak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, b, T)| specifying the tangent vector type%
%\footnote{the in-place variant \texttt{get\_vector!($\frak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
.
Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|%
%\footnote{in-place variant \texttt{get\_coordinates!($\frak g$, c, X, b)}}
, where the Basis is again optional.

For the case of the default, the \verb|DefaultLieAlgebraOrthogonalBasis()| the more commonly used names \verb|hat(|$\frak g$\verb|, c)|
%\footnote{in-place variant \texttt{hat!($\frak g$, X::T, c)}}
(again with an optional vector type \verb|T|) and \verb|vee(|$\frak g$\verb|, X)|
%\footnote{in-place variant \texttt{vee!($\frak g$, c, X)}}
are implemented

\subsubsection*{Push forward and pull back of tangent vectors}
To the representation of a tangent vector $X \in \frak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$. This is implemented as \verb|push_forward_tangent(G, g, X)|%
%\footnote{in-place variant \texttt{push\_forward\_tangent!(G, Y, g, X)}.}
.
Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\frak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$. This is implemented as \verb|pull_back_tangent(G, g, Y)|%
%\footnote{in-place variant \texttt{pull\_back\_tangent!(G, X, g, Y)}.}
.

@@

The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing funtionality.
For Lie groups there are currently two such decorators available

>>>>

\newcommand{\term}[1]{\emph{#1}}
\newcommand{\e}{\ensuremath{\mathrm{e}}}
\newcommand{\im}{\ensuremath{\mathrm{i}}}
\newcommand{\LieGroupsVersion}{0.1.8}
%
% Define muted colours.
\definecolor{TolMutedBlue}{HTML}{332288}
\definecolor{TolMutedCyan}{HTML}{88CCEE}
\definecolor{TolMutedTeal}{HTML}{44AA99}
\definecolor{TolMutedGreen}{HTML}{117733}
\definecolor{TolMutedOlive}{HTML}{999933}
\definecolor{TolMutedSand}{HTML}{DDCC77}
\definecolor{TolMutedRose}{HTML}{CC6677}
\definecolor{TolMutedWine}{HTML}{882255}
\definecolor{TolMutedPurple}{HTML}{AA4499}
%
\hypersetup{colorlinks=true}
%
\begin{document}

@@

These spaces are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
They have applications in physics, robotics, stochastic processes, information geometry, and many other areas see~\cite{Chirikjian:2009,Chirikjian:2012},
but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
and their numerical aspects, for example when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998,IserlesMunthe-KaasNoersettZanna:2000}.

@@

This paper provides an overview of the main features of \verb|LieGroups.jl| as of version \LieGroupsVersion{}, whose logo is shown on the right.
%
After introducing some mathematical background in Section~\ref{sec:Notation},
we present the interface in Section~\ref{sec:Interface} for the Lie group, the Lie algebra, and group actions.
Section~\ref{sec:LieGroups} provides an overview of all currently implemented Lie groups.
Finally, in Section~\ref{sec:Example}, we demonstrate how to get started and use \verb|LieGroups.jl|.

@@

which locally looks like \(\mathbb{R}^2\), think of charts in an atlas, but globally it is not.
Finally we denote the tangent space at a point \(p \in \mathcal{M}\) by \(T_p\mathcal{M}\). This can be thought of as all “velocities” (direction and speed) in which a curve can “pass through” a point. Formally it is set the equivalence classes of derivatives of smooth curves. Each tangent space $T_p\mathcal M$ is a $n$-dimensional vector space and we call the disjoint union of all tangent spaces
\begin{equation*}
    T\mathcal M = \dot\bigcup_{p \in \mathcal M} T_p\mathcal M
\end{equation*}
the \term{tangent bundle} of $\mathcal M$.
\\
A group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\). Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is, for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{Note that the identification is not continuous.}, or, continuously, with the circle. % chktex 9ok
The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose $p^{\mathrm{T}}$ (or an angle $-\alpha$). % -(-\pi) would be out of the range

\subsection{Tangent spaces and the Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
For a map \(f\colon \mathcal G \to \mathcal H\) between two Lie groups, its \term{differential} (or push forward) at a point \(g \in \mathcal G\) is a linear map \(Df(g)\colon T_g\mathcal G \to T_{f(g)}\mathcal H\) between the corresponding tangent spaces.
Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields. \texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
\\
As a first example, consider $\mathcal G = (\mathbb R^n, +)$, where the tangent space at any point $g$ is again $\mathbb R^n$, especially at the identity $e=0$. We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
Here, a tangent vector $X$ induces the constraint vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$. Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \mathfrak{g}$.

@@

Concerning a metric on the tangent spaces, smooth manifolds are turned into Riemannian manifolds when they are equipped with a Riemannian metric \(\langle \cdot, \cdot \rangle_p\colon T_p\mathcal{M} \times T_p\mathcal{M} \to \mathbb{R}\) for each point \(p \in \mathcal{M}\) that smoothly varies with \(p\).
For Lie groups, we can use a single inner product \(\langle \cdot, \cdot \rangle\) on the Lie algebra \(\mathfrak{g}\) and use the change in representation as mentioned above to define
\begin{equation*}
\langle X, Y \rangle_g = \langle D\lambda_{g^{-1}}(g)[X], D\lambda_{g^{-1}}(g)[Y] \rangle_e
\end{equation*}
for all \(X, Y \in T_g\mathcal{G}\) and \(g \in \mathcal{G}\). Representing tangent vectors at arbitrary points \(g \in \mathcal{G}\) using the Lie algebra \(\mathfrak{g}\) yields here, that we can use the single inner product directly to evaluate this Riemannian metric. Hence representing tangent vectors using the Lie algebra is the default in the following.
\\
For the two examples above we obtain the Euclidean inner product \(\langle X, Y \rangle = X^{\mathrm{T}} Y\) for \(X, Y \in \mathbb{R}^n\) for \((\mathbb{R}^n, +)\) and the Frobenius inner product \(\langle X, Y \rangle = \mathrm{trace}(X^{\mathrm{T}} Y)\), \(X, Y \in \mathfrak{so}(n)\) can be used for \(\mathrm{SO}(n)\).

@@

\subsection{Group Actions}\label{subsec:GroupActions}
A \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\sigma\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\sigma(e, p) = p\) and \(\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)\).\footnote{This is the convention for left actions. Alternatively, right actions fulfil \(\sigma(g, \sigma(h, p)) = \sigma(h \cdot g, p)\).}
Informally a group action describes how elements of the Lie group \(\mathcal{G}\) “act on” points on the manifold \(\mathcal{M}\). As an example, think of the special orthogonal group \(\mathrm{SO}(3)\) acting on points on Euclidean space $\mathbb{R}^3$ “moving” them somewhere by rotating them around the origin. We obtain the group action \(\sigma\colon \mathrm{SO}(3) \times \mathbb{R}^3 \to \mathbb{R}^3\), \(\sigma(R, x) \coloneqq Rx\).

The same action can also be applied to points from the sphere \(\mathbb{S}^2\), resulting in a similar group action \(\sigma\colon \mathrm{SO}(3) \times \mathbb{S}^2 \to \mathbb{S}^2\), \(\sigma(R, p) \coloneqq Rp\).

@@

The interface follows the philosophy of \verb|ManifoldsBase.jl|, that the Lie group or algebra is the first argument and even a mutated argument comes second.
Similarly, if a function computes something like a new point on the Lie group or a tangent vector, there also exists a variant, that computes this in-place.
For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default representation, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types. To generate a point in a specific representation of type \verb|T|, one can use \verb|identity_element(G, T)|.

@@

In general, when a function within the interfaces of \verb|JuliaManifolds| returns a point, a tangent vector or any other mutable object, there always exist two variants of the function, one that allocates a new object and returns it, and one that takes an additional argument to write the result into. For example for a function \verb|f(G, g)| that returns a point on the Lie group \verb|G|, there also exists a variant \verb|f!(G, h, g)| that computes the result in-place of \verb|h|, possibly using the memory also for interims results. All methods are written such that the result is correct even if arguments \verb|h| and \verb|g| use the same memory. By default the allocating variant is implemented by allocating memory accordingly and calling the in-place variant. This default can be overridden, for example for performance reasons.

\subsubsection*{Topological functions}
Points and tangent vectors can be represented by the same Julia objects, although they are distinct from a topological viewpoint.%
\footnote{\texttt{LieGroups.jl} introduces an abstract type \texttt{AbstractLieGroupPoint <: AbstractManifoldPoint}, that is used to represent points on the Lie group.
This abstract supertype is not necessary, points can be represented by arrays or other types as well.}
Separate types are necessary when one Lie group has \emph{different representations} of points, that have to be distinguished. When there are different representations, it is recommended to introduce a point type for each representation and make the default one fall back to (just) using arrays.
Several functions in the interface return points or tangent vectors, where the type might not be known from the input argument. For such a function, for example again a function \verb|f(G)| a positional argument \verb|T| can be used to specify the desired return type, i.e.\ \verb|f(G, T)| returns a point or tangent vector of type \verb|T|.
The in-place variant does not require this, since in its signature of the form \verb|f!(G, g::T)| the type \verb|T| is known from the argument \verb|g|.

@@

In \verb|LieGroups.jl|, the group exponential function is given by \verb|exp(G,X)|%
%\footnote{in-place variant \texttt{exp!(G, g, X)}.}
.
If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\mathfrak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|%
%\footnote{in-place variant \texttt{exp!(G, h, g, X)}.}
.
As a word of carefulness, note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map. The Riemannian exponential map \verb|exp(M, p, X)|
%\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.}
can be distinguished in that the first argument is a manifold, and the tangent vector \verb|X| has to be from the tangent space at the point \verb|p| on the manifold.
Here, again, to access the Riemannian exponential, one can use the base manifold of the Lie group, i.e.\ \verb|exp(base_manifold(G), g, X)| and have to make sure that \verb|X| is from the tangent space at \verb|g|.

@@

Similar to points on the Lie group, when representing elements $X \in \mathfrak g$ from a Lie algebra, we do not type the general functions of the interface.
This allows to use either plain arrays or own structures to represent these in code.
\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}
Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.

@@

A central function on the Lie algebra is the Lie bracket $[\cdot, \cdot]\colon \mathfrak g \times \mathfrak g \to \mathfrak g$, which is available as \verb|lie_bracket(|$\mathfrak g$\verb|, X, Y)|.
% \footnote{in-place variant \texttt{lie\_bracket!($\mathfrak g$, Z, X, Y)}.}
For more details on the Lie bracket see~\cite[Ch.~5]{HilgertNeeb:2012}.

\subsubsection*{Topological basics.}
Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\mathfrak g$ \verb|= LieAlgebra(G)|.
To access the Lie group again, use \verb|base_lie_group(|$\mathfrak g$\verb|)|. Similarly \verb|base_manifold(|$\mathfrak g$\verb|)| returns the underlying manifold of the Lie group.\\
As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
A zero vector is generated via \verb|zero_vector(|$\mathfrak g$\verb|)| for the default representation and \verb|zero_vector(|$\mathfrak g$\verb|, T)| for a specific representation type \verb|T|%
%\footnote{the in-place variant \texttt{zero\_vector!($\mathfrak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
.

The main topological function is to test the validity of a tangent vector \verb|is_vector(|$\mathfrak g$\verb|, X; error=:none)| using the same \verb|error=| keyword as \verb|is_point| on the Lie group.

\subsubsection*{Vector space related functions}
For the following functions related to vector space features, using the Lie algebra $\mathfrak g$ as first argument is equivalent to specifying the Lie Group $G$ and an arbitrary point $g \in \mathcal G$. This yields that the Lie group complies with the general interface for manifolds.

The inner product and norm on the Lie algebra are available as \verb|inner(|$\mathfrak g$\verb|, X, Y)| and \verb|norm(|$\mathfrak g$\verb|, X)|, respectively.
Furthermore, there are two functions to convert between a coordinate-free representation of $X$ as a tangent vector and its representation in coordinates of a basis.

Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b, T)| specifying the tangent vector type%
%\footnote{the in-place variant \texttt{get\_vector!($\mathfrak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
.
Given a tangent vector $X \in \mathfrak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\mathfrak g$\verb|, X, B)|%
%\footnote{in-place variant \texttt{get\_coordinates!($\mathfrak g$, c, X, b)}}
, where the Basis is again optional.

For the case of the default, the \verb|DefaultLieAlgebraOrthogonalBasis()| the more commonly used names \verb|hat(|$\mathfrak g$\verb|, c)|
%\footnote{in-place variant \texttt{hat!($\mathfrak g$, X::T, c)}}
(again with an optional vector type \verb|T|) and \verb|vee(|$\mathfrak g$\verb|, X)|
%\footnote{in-place variant \texttt{vee!($\mathfrak g$, c, X)}}
are implemented

\subsubsection*{Push forward and pull back of tangent vectors}
To the representation of a tangent vector $X \in \mathfrak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$. This is implemented as \verb|push_forward_tangent(G, g, X)|%
%\footnote{in-place variant \texttt{push\_forward\_tangent!(G, Y, g, X)}.}
.
Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\mathfrak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$. This is implemented as \verb|pull_back_tangent(G, g, Y)|%
%\footnote{in-place variant \texttt{pull\_back\_tangent!(G, X, g, Y)}.}
.

@@

The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing functionality.
For Lie groups there are currently two such decorators available
